---
title: "prosperity"
author: "Xuhui Ying, Zach Ahearn, Pen Trainer, Jenny Yue, Leo Li"
date: "09/03/2022"
output:
  pdf_document:
    toc: yes
    toc_depth: '3'
  html_document:
    toc: yes
    toc_depth: 3
    theme: paper
    highlight: tango
    df_print: paged
---

# load libraries

```{r, eval=TRUE, message=FALSE, warning=FALSE}
library(ggplot2)
library(corrplot)
library(MASS)
library(skimr)
library(readr)
#library(VIF)
#library(VIFCP)
library(fastDummies)
library(tidyverse)
library(janitor)
library(modelr)

options(scipen = 999) # turns off scientific notation
```

# load data

```{r, eval=TRUE, message=FALSE, warning=FALSE}
df <- read_csv(file="WorldBank2002_2.csv") %>% clean_names()

head(df)

#data <- df %>% dplyr::select(-1)
data <- df %>% select_if(is.numeric)

skim(data)

#Missing Values
skim(data)
nrow(data)

#Missing Value Percentages by Column
missing_values <- colMeans(is.na(data)*100)
missing_values


#Missing Value Percentages by Row
rowSums(is.na(data))/47*100


prosperity <- df %>%
  dplyr::select(-military_expenditure_percent_of_gdp, -personal_computers_per_1_000_p) %>%
  filter(series != "Somalia") %>% #Intially Just Somalia but then added 3 more later due to outliers
  filter(series != "Angola") %>%
  filter(series != "South Africa") %>%
  filter(series != "Zimbabwe") %>% 
  select_if(is.numeric)

#prosperity %>% select_if(is.numeric) %>% skim()

prosperity$co2_emissions_metric_tons_per[is.na(prosperity$co2_emissions_metric_tons_per)]<-median(prosperity$co2_emissions_metric_tons_per,na.rm=TRUE)
prosperity$exports_of_goods_and_services[is.na(prosperity$exports_of_goods_and_services)]<-median(prosperity$exports_of_goods_and_services,na.rm=TRUE)
prosperity$fertility_rate_total_births_p[is.na(prosperity$fertility_rate_total_births_p)]<-median(prosperity$fertility_rate_total_births_p,na.rm=TRUE)

prosperity$fixed_line_and_mobile_phone_sub[is.na(prosperity$fixed_line_and_mobile_phone_sub)]<-median(prosperity$fixed_line_and_mobile_phone_sub,na.rm=TRUE)
prosperity$foreign_direct_investment_net[is.na(prosperity$foreign_direct_investment_net)]<-median(prosperity$foreign_direct_investment_net,na.rm=TRUE)
prosperity$gni_per_capita_atlas_method_c[is.na(prosperity$gni_per_capita_atlas_method_c)]<-median(prosperity$gni_per_capita_atlas_method_c,na.rm=TRUE)



prosperity$gni_atlas_method_current_us[is.na(prosperity$gni_atlas_method_current_us)]<-median(prosperity$gni_atlas_method_current_us,na.rm=TRUE)
prosperity$gross_capital_formation_percent_of_g[is.na(prosperity$gross_capital_formation_percent_of_g)]<-median(prosperity$gross_capital_formation_percent_of_g,na.rm=TRUE)
prosperity$imports_of_goods_and_services[is.na(prosperity$imports_of_goods_and_services)]<-median(prosperity$imports_of_goods_and_services,na.rm=TRUE)

prosperity$internet_users_per_1_000_peopl[is.na(prosperity$internet_users_per_1_000_peopl)]<-median(prosperity$internet_users_per_1_000_peopl,na.rm=TRUE)
prosperity$long_term_debt_dod_current_us[is.na(prosperity$long_term_debt_dod_current_us)]<-median(prosperity$long_term_debt_dod_current_us,na.rm=TRUE)
prosperity$net_barter_terms_of_trade_2000[is.na(prosperity$net_barter_terms_of_trade_2000)]<-median(prosperity$net_barter_terms_of_trade_2000,na.rm=TRUE)

prosperity$short_term_debt_outstanding_do[is.na(prosperity$short_term_debt_outstanding_do)]<-median(prosperity$short_term_debt_outstanding_do,na.rm=TRUE)
prosperity$ln_gni[is.na(prosperity$ln_gni)]<-median(prosperity$ln_gni,na.rm=TRUE)


```
# Full Model

```{r, eval=TRUE, message=FALSE, warning=FALSE}

#Fit multiple regression model

fullreg <- lm(gdp_current_us~., prosperity)

fullreg
summary(fullreg)
plot(fullreg)

#check for multicollinearity

vifres_1 <- car::vif(fullreg)
vifres_1

```
# Reduced Model - backward elimination (stepAIC)

```{r, eval=TRUE, message=FALSE, warning=FALSE}

# Try Stepwise Selection to reduce model

step <- stepAIC(fullreg, direction="both")
summary(step)
plot(step)

# check reduced model for multicollinearity

vifres_2 <- car::vif(step)
vifres_2

```

# Reduced Model - Remove ln_gni (p-value > 0.05)

```{r, eval=TRUE, message=FALSE, warning=FALSE}

model_3 <- lm(gdp_current_us ~ agriculture_value_added_percent_of + 
    co2_emissions_metric_tons_per + fertility_rate_total_births_p + 
    foreign_direct_investment_net + gni_atlas_method_current_us + 
    gross_capital_formation_percent_of_g + immunization_measles_percent_of_chi + 
    imports_of_goods_and_services + inflation_gdp_deflator_annual + 
    internet_users_per_1_000_peopl + long_term_debt_dod_current_us + 
    merchandise_trade_percent_of_gdp + official_development_assistance + 
    population_total + services_etc_value_added_percent + surface_area_sq_km, data = prosperity)

model_3
summary(model_3)
plot(model_3)

#check for multicollinearity

vifres_3 <- car::vif(model_3)
vifres_3

```

# Reduced Model - Remove agriculture_value_added_percent_of (p-value > 0.05)

```{r, eval=TRUE, message=FALSE, warning=FALSE}

model_4 <- lm(gdp_current_us ~  co2_emissions_metric_tons_per + fertility_rate_total_births_p + 
    foreign_direct_investment_net + gni_atlas_method_current_us + 
    gross_capital_formation_percent_of_g + immunization_measles_percent_of_chi + 
    imports_of_goods_and_services + inflation_gdp_deflator_annual + 
    internet_users_per_1_000_peopl + long_term_debt_dod_current_us + 
    merchandise_trade_percent_of_gdp + official_development_assistance + 
    population_total + services_etc_value_added_percent + surface_area_sq_km, data = prosperity)

model_4
summary(model_4)
plot(model_4)

#check for multicollinearity

vifres_4 <- car::vif(model_4)
vifres_4

```

# Reduced Model - Remove services_etc_value_added_percent (p-value > 0.05)

```{r, eval=TRUE, message=FALSE, warning=FALSE}

model_5 <- lm(gdp_current_us ~  co2_emissions_metric_tons_per + fertility_rate_total_births_p + 
    foreign_direct_investment_net + gni_atlas_method_current_us + 
    gross_capital_formation_percent_of_g + immunization_measles_percent_of_chi + 
    imports_of_goods_and_services + inflation_gdp_deflator_annual + 
    internet_users_per_1_000_peopl + long_term_debt_dod_current_us + 
    merchandise_trade_percent_of_gdp + official_development_assistance + 
    population_total + surface_area_sq_km, data = prosperity)

model_5
summary(model_5)
plot(model_5)

#check for multicollinearity

vifres_5 <- car::vif(model_5)
vifres_5

```

# Reduced Model - Remove gross_capital_formation_percent_of_g (p-value > 0.05)

```{r, eval=TRUE, message=FALSE, warning=FALSE}

model_6 <- lm(gdp_current_us ~  co2_emissions_metric_tons_per + fertility_rate_total_births_p + 
    foreign_direct_investment_net + gni_atlas_method_current_us + immunization_measles_percent_of_chi + 
    imports_of_goods_and_services + inflation_gdp_deflator_annual + 
    internet_users_per_1_000_peopl + long_term_debt_dod_current_us + 
    merchandise_trade_percent_of_gdp + official_development_assistance + 
    population_total + surface_area_sq_km, data = prosperity)

model_6
summary(model_6)
plot(model_6)

#check for multicollinearity

vifres_6 <- car::vif(model_6)
vifres_6

```

# Reduced Model - Remove internet_users_per_1_000_peopl (p-value > 0.05)

```{r, eval=TRUE, message=FALSE, warning=FALSE}

model_7 <- lm(gdp_current_us ~ co2_emissions_metric_tons_per + fertility_rate_total_births_p + 
    foreign_direct_investment_net + gni_atlas_method_current_us + immunization_measles_percent_of_chi + 
    imports_of_goods_and_services + inflation_gdp_deflator_annual + long_term_debt_dod_current_us + 
    merchandise_trade_percent_of_gdp + official_development_assistance + 
    population_total + surface_area_sq_km, data = prosperity)

model_7
summary(model_7)
plot(model_7)

#check for multicollinearity

vifres_7 <- car::vif(model_7)
vifres_7

```

# Reduced Model - Remove gni_atlas_method_current_us (VIF > 10)

```{r, eval=TRUE, message=FALSE, warning=FALSE}

model_8 <- lm(gdp_current_us ~ co2_emissions_metric_tons_per + fertility_rate_total_births_p + 
    foreign_direct_investment_net + immunization_measles_percent_of_chi + 
    imports_of_goods_and_services + inflation_gdp_deflator_annual + long_term_debt_dod_current_us + 
    merchandise_trade_percent_of_gdp + official_development_assistance + 
    population_total + surface_area_sq_km, data = prosperity)

model_8
summary(model_8)
plot(model_8)

#check for multicollinearity

vifres_8 <- car::vif(model_8)
vifres_8

```

# Reduced Model - Remove gni_atlas_method_current_us (VIF > 5)

```{r, eval=TRUE, message=FALSE, warning=FALSE}

model_9 <- lm(gdp_current_us ~ co2_emissions_metric_tons_per + fertility_rate_total_births_p + 
    foreign_direct_investment_net + immunization_measles_percent_of_chi + 
    imports_of_goods_and_services + inflation_gdp_deflator_annual + merchandise_trade_percent_of_gdp + official_development_assistance + 
    population_total + surface_area_sq_km, data = prosperity)

model_9
summary(model_9)
plot(model_9)

#check for multicollinearity

vifres_9 <- car::vif(model_9)
vifres_9

```

# Reduced Model - Remove merchandise_trade_percent_of_gdp (p-value > 0.05)

```{r, eval=TRUE, message=FALSE, warning=FALSE}

model_10 <- lm(gdp_current_us ~ co2_emissions_metric_tons_per + fertility_rate_total_births_p + 
    foreign_direct_investment_net + immunization_measles_percent_of_chi + 
    imports_of_goods_and_services + inflation_gdp_deflator_annual + official_development_assistance + 
    population_total + surface_area_sq_km, data = prosperity)

model_10
summary(model_10)
plot(model_10)

#check for multicollinearity

vifres_10 <- car::vif(model_10)
vifres_10

```

# Reduced Model - Remove co2_emissions_metric_tons_per (p-value > 0.05)

```{r, eval=TRUE, message=FALSE, warning=FALSE}

model_11 <- lm(gdp_current_us ~ fertility_rate_total_births_p + 
    foreign_direct_investment_net + immunization_measles_percent_of_chi + 
    imports_of_goods_and_services + inflation_gdp_deflator_annual + official_development_assistance + 
    population_total + surface_area_sq_km, data = prosperity)

model_11
summary(model_11)
plot(model_11)

#check for multicollinearity

vifres_11 <- car::vif(model_11)
vifres_11

```

# Reduced Model - Remove official_development_assistance (p-value > 0.05)

```{r, eval=TRUE, message=FALSE, warning=FALSE}

model_12 <- lm(gdp_current_us ~ fertility_rate_total_births_p + 
    foreign_direct_investment_net + immunization_measles_percent_of_chi + 
    imports_of_goods_and_services + inflation_gdp_deflator_annual + population_total + surface_area_sq_km, data = prosperity)

model_12
summary(model_12)
plot(model_12)

#check for multicollinearity

vifres_12 <- car::vif(model_12)
vifres_12

```

# Reduced Model - Remove inflation_gdp_deflator_annual (p-value > 0.05)

```{r, eval=TRUE, message=FALSE, warning=FALSE}

model_13 <- lm(gdp_current_us ~ fertility_rate_total_births_p + 
    foreign_direct_investment_net + immunization_measles_percent_of_chi + 
    imports_of_goods_and_services + population_total + surface_area_sq_km, data = prosperity)

model_13
summary(model_13)
plot(model_13)

#check for multicollinearity

vifres_13 <- car::vif(model_13)
vifres_13

```

# Reduced Model - Remove immunization_measles_percent_of_chi (p-value > 0.05)

```{r, eval=TRUE, message=FALSE, warning=FALSE}

model_14 <- lm(gdp_current_us ~ fertility_rate_total_births_p + 
    foreign_direct_investment_net + imports_of_goods_and_services + population_total + surface_area_sq_km, data = prosperity)

model_14
summary(model_14)
plot(model_14)

#check for multicollinearity

vifres_14 <- car::vif(model_14)
vifres_14

```

# Reduced Model - take a log transformation

```{r, eval=TRUE, message=FALSE, warning=FALSE}

model_15 <- lm(log10(gdp_current_us) ~ fertility_rate_total_births_p + 
    foreign_direct_investment_net + imports_of_goods_and_services + log10(population_total) + surface_area_sq_km, data = prosperity)

model_15
summary(model_15)
plot(model_15)

#check for multicollinearity

vifres_15 <- car::vif(model_15)
vifres_15

```

# predictions

```{r}
data_predict <- add_predictions(prosperity, var = "winnings_pred", model_15)

residual <- data_predict %>%
  mutate(compare = if_else(winnings_pred<log10(gdp_current_us), 'underestimated', 'overestimated'), error = log10(gdp_current_us) - winnings_pred, abs_error = abs(log10(gdp_current_us) - winnings_pred),log_gdp = log10(gdp_current_us)) %>%
  dplyr::select(log_gdp, winnings_pred, compare, error, abs_error)

residual

residual %>%
  slice_min(order_by = abs_error, n = 10) %>%
  dplyr::select(log_gdp, winnings_pred, compare, abs_error)

residual %>%
  slice_min(order_by = error, n = 10) %>%
  dplyr::select(log_gdp, winnings_pred, compare, error)
  
residual %>%
  slice_max(order_by = error, n = 10) %>%
  dplyr::select(log_gdp, winnings_pred, compare, error)
```
